Updated Ephemerides of the young multi-planet system V1298 Tau from TESS

### 2021-10-19 src/ms.tex
updates

% Define document class
\documentclass[twocolumn]{aastex631}

% Filler text
\include{style}

% Begin!
\begin{document}

% Title
\title{Updated Ephemerides of the young multi-planet system V1298 Tau from TESS}

% Author list
\author[0000-0002-9464-8101]{Adina~D.~Feinstein}
\altaffiliation{NSF Graduate Research Fellow}
\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}

@@

% Abstract with filler text
\begin{abstract}
  \sname is a young ($\sim$23~Myr) solar analogue hosting four transiting exoplanets. 
  Given its planets range between 0.5 - 0.9 $R_J$, this system provides a unique opportunity to 
  understand the evolution of planetary radii in the same stellar environment. \sname was originally 
  observed 6 years ago during K2 Campaign 4. The extended mission of NASA's Transiting Exoplanet Survey 
  Satellite (\tess) includes observing the ecliptic plane. Here, we present new photometric observations 
  of \sname from the 10-minute TESS Full-Frame Images. We use the TESS data to update the transit-timing 
  for \allplanets as well as compare newly observed radii to that derived from the K2 light curve. We 
  additionally catch a transit of \planete, which was a single-transit event in the previous data, allowing 
  us to further constrain the period of this farthest known planet. \sname is the target of several ongoing 
  and future observations, including with the James Webb Space Telescope. These updated ephemerides will be 
  crucial for accurately recovering transit events and understanding any future doppler tomographic or 
  transmission spectroscopy signals.

\end{abstract}

@@

% Main body with filler text
\section{Introduction}

Planetary sizes are expected to evolve over time, due to a variety of endogenous and exogenous physical processes including gravitational contraction, atmospheric heating and mass-loss, and core-envelope interactions \citep[e.g.][]{OwenWu2013, Lopez2013, Jin2014, ChenRogers2016, Ginzburg2018}. Dramatic size evolution is expected at early stages, when planets are still contracting and radiating away the energy from their formation, and when host stars are heating planetary atmospheres with high levels of X-ray and ultraviolet radiation. Since the size evolution of any individual planet is believed to be slow relative to typical observational baselines, the best way to make inferences about the size evolution of exoplanets is by measuring the sizes of large numbers of planets across a range of ages.

@@

@@

During its Extended Mission Cycle 4, \tess\ is re-observing many of the previous \textit{K2} fields. \sname (TIC 15756231) was observed by \tess\ in Sectors 43 (UT 16 Sep 2021 to UT 12 Oct 2021) at 10-minute cadence within the Full-Frame Images (FFIs) and at 2-minute cadence, which are processed by the Science Processing Operations Center (SPOC) pipeline \citep{jenkinsSPOC2016}.

Initially, we used the \texttt{tica} \citep{fausnaugh20} software to download calibrated FFIs each orbit for Sectors 43 and 44. These FFIs are quickly available after the data is down-linked. We created light curves from the \texttt{tica}-processed FFIs by modeling the point-spread function (PSF) of \sname. Due to crowding of two other nearby bright sources (see Figure~\ref{fig:tpf}), we found that modeling each star (3 in total) with a 2D Gaussian was sufficient for extracting a non-contaminated light curve. %This allowed us to place initial constraints on the planets' ephemerides. The remaining analysis was completed using the SPOC-processed light curves.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth,trim={0.25cm 0 0 0}]{static/TESSaperture.pdf}
\caption{The \tess\ target pixel file (TPF) overlaid with a sky image of \sname taken with the Sloan Digital Sky Survey (SDSS) i-band. \sname is shown in the white circle; nearby sources with \tess\ magnitudes $< 14$ are marked with orange x's. There are two bight nearby sources to \sname. However, selecting the one pixel aperture (white square) encapsulates only \sname and therefore contamination from the nearby sources is negligible. \label{fig:tpf} }
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht]
\begin{center}
\includegraphics[width=\textwidth,trim={0.25cm 0 0 0}]{static/lightcurve.pdf}
\caption{\sname extracted light curve from the \texttt{tica}-processed full-frame images, with transits of \allplanets highlighted by color. Top row: extracted light curve with over plotted with our best-fit GP model for stellar variability (black). Middle row: the \tess\ light curve with the stellar variability removed by our model. Bottom rows: zoomed-in regions around the visible transits during the first orbit (third row) and second orbit (last row) from \tess\ Sector 43. \label{fig:transits}}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We simultaneously modeled the transits of \allplanets and the stellar variability using the open-source packages \exoplanet \citep{exoplanet2019, exoplanet2021} and \texttt{PyMC3} \citep{Salvatier16}. Transit timings were originally identified using updated ephemerides from \textit{Spitzer} (Livingston et al. in prep). All other transit parameters (presented in Table~\ref{tab:table}) were initialized using values from \cite{David2019a}. We assumed quadratic limb darkening law, following the reparameterization described by \cite{kipping13}; this method allows for an efficient and uniform sampling of limb-darkening fits.

@@

\subsection{\textit{K2} vs. \textit{TESS} Radii}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.465\textwidth,trim={0.25cm 0 4.5cm 0}]{static/folded_compare.pdf}
\caption{Phase-folded \tess\ data (gray) with the new best-fit model (color) compared to the best-fit parameters from the original \textit{K2} data (black). The residuals between the data and each fit are shown underneath. The transit depth from the \textit{K2} data is deeper than that from \tess, particularly for \planetb and \planete. \label{fig:compare}}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

A full transit-timing variation (TTV) analysis is presented in Livingston et al.

@@

\begin{deluxetable*}{l r r r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{\sname light curve fitting results. \label{tab:table}}
\tablehead{\\
\hline\
\textit{Star} & \textit{Value} & \textit{Prior} &  & \\
\hline\
$M_\star [M_\odot]$ & & & & \\
$R_\star [R_\odot]$ & & & & \\
$u_1$ & & & & \\
$u_2$ & & & & \\
$P_{rot}$ [days] & & & & \\
ln($A$ [ppt]) & & & & \\
ln($Q_0$) & & & & \\
$\Delta Q_0$ & & & & \\
f [ppt] & & & & \\
mix & & & & \\
\hline\
\textit{Planets} & \textit{c} & \textit{d} & \textit{b} & \textit{e}\\
\hline\
P [days]               &  &  &  &  \\
$T_0$ [BJD - 2457000]  &  &  &  &  \\
Depth [ppt]            &  &  &  &  \\
$R_p/R_\star$          &  &  &  &  \\
$a/R_\star$            &  &  &  &  \\
$b$                    &  &  &  &  \\
$e$ [$^\circ$]         &  &  &  &  \\
\hline\
\textit{Priors} & \textit{c} & \textit{d} & \textit{b} & \textit{e}\\
\hline\
log($P$ [days]) & & & & \\
$T_0$ [BJD - 2457000] & & & & \\
log($R_p/R_\oplus$) & & & & \\
$b$ & & & & \\
$e$ & & & & \\
$\omega$ [$^\circ$] & & & & 
}
\startdata
\enddata
%\tablecomments{}
\end{deluxetable*}

@@

>>>>

\documentclass[twocolumn]{aastex631}

\include{style}

\begin{document}

\title{Updated Ephemerides of the young multi-planet system V1298 Tau from TESS}

\author[0000-0002-9464-8101]{Adina~D.~Feinstein}
\altaffiliation{NSF Graduate Research Fellow}
\affiliation{Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA}

@@

%%%%%%%%%%%%%%%%%%%%
% abstract must be < 250 words
% currently at 180 words
%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\sname is a young ($\sim$23~Myr) solar analogue hosting four transiting exoplanets. Given its planets range between 0.5 - 0.9 $R_J$, this system provides a unique opportunity to understand the evolution of planetary radii in the same stellar environment. \sname was originally observed 6 years ago during K2 Campaign 4. The extended mission of NASA's Transiting Exoplanet Survey Satellite (\tess) includes observing the ecliptic plane. Here, we present new photometric observations of \sname from the 10-minute TESS Full-Frame Images. We use the TESS data to update the transit-timing for \allplanets as well as compare newly observed radii to that derived from the K2 light curve. We additionally catch a transit of \planete, which was a single-transit event in the previous data, allowing us to further constrain the period of this farthest known planet. \sname is the target of several ongoing and future observations, including with the James Webb Space Telescope. These updated ephemerides will be crucial for accurately recovering transit events and understanding any future doppler tomographic or transmission spectroscopy signals.\end{abstract}

%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro}
Planetary sizes are expected to evolve over time, due to a variety of endogenous and exogenous physical processes including gravitational contraction, atmospheric heating and mass-loss, and core-envelope interactions \citep[e.g.][]{OwenWu2013, Lopez2013, Jin2014, ChenRogers2016, Ginzburg2018}. Dramatic size evolution is expected at early stages, when planets are still contracting and radiating away the energy from their formation, and when host stars are heating planetary atmospheres with high levels of X-ray and ultraviolet radiation. Since the size evolution of any individual planet is believed to be slow relative to typical observational baselines, the best way to make inferences about the size evolution of exoplanets is by measuring the sizes of large numbers of planets across a range of ages.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.46\textwidth,trim={0.25cm 0 0 0}]{static/TESSaperture.pdf}
\caption{The \tess\ target pixel file (TPF) overlaid with a sky image of \sname taken with the Sloan Digital Sky Survey (SDSS) i-band. \sname is shown in the white circle; nearby sources with \tess\ magnitudes $< 14$ are marked with orange x's. There are two bight nearby sources to \sname. However, selecting the one pixel aperture (white square) encapsulates only \sname and therefore contamination from the nearby sources is negligible. \label{fig:tpf}}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

During its Extended Mission Cycle 4, \tess\ is re-observing many of the previous \textit{K2} fields. \sname (TIC 15756231) was observed by \tess\ in Sectors 43 (UT 16 Sep 2021 to UT 12 Oct 2021) at 10-minute cadence within the Full-Frame Images (FFIs). We used the \texttt{tica} \citep{fausnaugh20} software to download calibrated FFIs each orbit for Sectors 43, as these FFIs are quickly available after the data is down-linked. We created light curves from the \texttt{tica}-processed FFIs by modeling the point-spread function (PSF) of \sname and the two nearby bright sources (see Figure~\ref{fig:tpf}). While aperture photometry provided a decent light curve, we found that modeling each star (3 in total) with a 2D Gaussian created a less contaminated light curve for \sname (Figure~\ref{fig:transits}, top row). %This allowed us to place initial constraints on the planets' ephemerides. The remaining analysis was completed using the SPOC-processed light curves.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!ht]
\begin{center}
\includegraphics[width=\textwidth,trim={0.25cm 0 0 0}]{static/lightcurve.pdf}
\caption{\sname extracted light curve from the \texttt{tica}-processed full-frame images, with transits of \allplanets highlighted by color. Top row: extracted light curve with over plotted with our best-fit GP model for stellar variability (blue). Middle row: the \tess\ light curve with the stellar variability removed by our model. Bottom rows: zoomed-in regions around the visible transits during the first orbit (third row) and second orbit (last row) from \tess\ Sector 43. \label{fig:transits}}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.465\textwidth,trim={0.25cm 0 4.5cm 0}]{static/folded_compare.pdf}
\caption{Phase-folded \tess\ data (gray) with the new best-fit model (color) compared to the best-fit parameters from the original \textit{K2} data (black). The residuals between the data and each fit are shown underneath. A comparison of measured $R_p/R_\star$ are presented at the bottom, with a 1-to-1 line plotted in black for reference. The transit depth from the \textit{K2} data is deeper than that from \tess, particularly for \planetb and \planete. \label{fig:compare}}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% planet fits
We simultaneously modeled the transits of \allplanets and the stellar variability using the open-source packages \exoplanet \citep{exoplanet2019, exoplanet2021} and \texttt{PyMC3} \citep{Salvatier16}. Transit timings were originally identified using updated ephemerides from \textit{Spitzer} (Livingston et al. in prep). All other transit parameters (presented in Table~\ref{tab:table}) were initialized using values from \cite{David2019a}. We assumed quadratic limb darkening law, following the reparameterization described by \cite{kipping13}; this method allows for an efficient and uniform sampling of limb-darkening fits.

@@

\subsection{\textit{K2} vs. \tess}

We compare the differences in transit parameters between the \textit{K2} and \tess\ data in Figure~\ref{fig:compare}. We plot the best-fit transit model from \textit{K2} in black, which was generated using parameters from \cite{David2019a} and the \texttt{batman} transit-modeling software \citep{Kreidberg15}. It is important to note that the \textit{K2} bandpass ranges from $\sim 400-900$~nm \citep{Howell2014}, while the \tess\ bandpass ranges from $\sim 600-1000$~nm \citep{Ricker2015}.

@@

There was a single transit of \planete in the original \textit{K2} data, which occurred roughly in the middle of the campaign. Since no other transits were detected, this provides a lower period limit of 36~days.

@@

We present new ephemerides for all four known planets in the \sname system. We have recovered the second transit of \planete; this new transit time in combination from the transit observed with \textit{K2} has allowed us to place tighter constraints on the period of the outermost planet.

A full transit-timing variation (TTV) analysis using \textit{K2}, \textit{Spitzer}, and \tess\ ephemerides is presented in Livingston et al. (in prep).

@@

\begin{deluxetable*}{l r r r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{\sname light curve fitting results. \label{tab:table}}
\tablehead{\\
\hline\
\textit{Star} & \textit{Value} & \textit{Prior} &  & \\
\hline
$R_\star [R_\odot]$ & & $\mathcal{G}(1.305, 0.07)$ & & \\
$u_1$ & & & & \\
$u_2$ & & & & \\
$P_{rot}$ [days] & & $\mathcal{G}($ln$ 2.87, 2)$ & & \\
ln($A$ [ppt]) & & & & \\
ln($Q_0$) & & $\mathcal{H}(\sigma=2)$ & & \\
$\Delta Q_0$ & & $\mathcal{G}(0, 2)$ & & \\
$\rho_\textrm{SHO}$ & & $\mathcal{G}(0, 10)$ & & \\
$\sigma_\textrm{SHO}$ & & $\mathcal{G}(\sigma_\textrm{flux}, 10)$ & & \\
f [ppt] & & & & \\
mix & & & & \\
\hline\
\textit{Light Curve} & \textit{Value} & \textit{Prior} &  & \\
\hline
trend term a & & $\mathcal{G}(0, 0.01)$ & & \\
trend term b & & $\mathcal{G}(0, 0.1)$ & & \\
trend term c & & $\mathcal{G}(0, 1)$ & & \\
$\sigma_l$ & & $\mathcal{G}($ln$(0.1\sigma_\textrm{flux}), 10)$  & & \\
$\sigma_j$ & & $\mathcal{G}($ln$(0.1\sigma_\textrm{flux}), 10)$  & & \\
\hline\
\textit{Planets} & \textit{c} & \textit{d} & \textit{b} & \textit{e}\\
\hline
P [days]               &  &  &  &  \\
$T_0$ [BJD - 2457000]  &  &  &  &  \\
Depth [ppt]            &  &  &  &  \\
$R_p/R_\star$          &  &  &  &  \\
$a/R_\star$            &  &  &  &  \\
$b$                    &  &  &  &  \\
$e$ [$^\circ$]         &  &  &  &  \\
\hline\
\textit{Priors} & \textit{c} & \textit{d} & \textit{b} & \textit{e}\\
\hline
log($P$ [days]) & $\mathcal{G}($ln$ 8.25, 1)$ & $\mathcal{G}($ln$ 12.40, 1)$& $\mathcal{G}($ln$ 24.14, 1)$ & $\mathcal{G}($ln$ 36.70, 1)$\\
$T_0$ [BJD - 2457000] & $\mathcal{G}(4648.53,0.1)$ & $\mathcal{G}(4645.4,0.1)$ & $\mathcal{G}(4648.1,0.1)$ & $\mathcal{G}(4648.8,0.1)$ \\
log(depth [ppt]) & $\mathcal{G}($ln$ 1.45, 1)$ & $\mathcal{G}($ln$ 1.90, 1)$ & $\mathcal{G}($ln$ 4.90, 1)$ & $\mathcal{G}($ln$ 3.73, 1)$ \\
$b$ & $\mathcal{U}[0, 1]$ & $\mathcal{U}[0, 1]$ & $\mathcal{U}[0, 1]$ & $\mathcal{U}[0, 1]$ \\
duration [days] & $\mathcal{G}($ln$ 0.19, 1)$ & $\mathcal{G}($ln$ 0.23, 1)$ & $\mathcal{G}($ln$ 0.27, 1)$ &  $\mathcal{G}($ln$ 0.31, 1)$
}
\startdata
\enddata
\tablecomments{Priors are noted for parameters that were directly sampled. The distributions are as follows -- $\mathcal{G}$: Gaussian; $\mathcal{H}$: Half-normal; $\mathcal{U}$: Uniform. $\sigma_\textrm{flux}$ is the standard deviation of the light curve}
\end{deluxetable*}

@@

\include{table}
